---
title: "Young children’s developing ability to integrate gestural and emotional cues"
date: "24 2 2018"
output:
  html_document: default
---

```{r load_packages, include = FALSE}
library("lme4")
library("ggplot2")
library("tidyverse")
library("ggthemes")
library("skimr")
library("lsr")
library("langcog")

# langcog package is installed by running
# install.packages("devtools")
# devtools::install_github("langcog/langcog")


#### loading data
data <- read.csv("data/emopoint.data.csv", sep=",")
```
# Data structure:

* $ study       : Study 1 - 3
* $ age         : Age group (22,28,34,40 or 46)
* $ phase       : Point or peek used as cue
* $ id          : Unique number identifying each subject
* $ aged        : Participant age in days
* $ trial       : Trial within participant
* $ type        : Emotional expression: positive or negative
* $ corr        : Correct choice
* $ sex         : 0 = female
* $ unique_group: Variable uniquely identifying each cell (combination of age, phase and type)
 
# Demographics summary
```{r participant ovrview, echo = FALSE}
data %>%
  filter (trial == "1") %>%
  group_by(study,phase,age, type) %>%
  arrange(study, age) %>%
  summarise(m_age = mean(aged)/30.42,
            sd_age = sd(aged)/30.42,
            ll_age = min(aged)/30.42,
            ul_age = max(aged)/30.42,
            sex_m = sum (sex)) %>%
  knitr::kable(digits = 2)
```

# Comparing performance to chance

Two-tailed one sample t-tests for group level comparison to chance. 
Individual performance is evaluated based on two-tailed binomial tests with p > .05 above chance when all trials are correct and below chance if all trials are incorrect.
```{r chance performance, echo = FALSE}
t <- data %>%
  group_by(study,phase,age, type, id) %>%
  arrange(study, age) %>%
  summarise(corr = mean(corr)) %>%
  summarise(corr = list(corr)) %>%
  group_by(study,phase,age, type) %>%
  mutate(m = mean(unlist(corr)),
         sd = sd(unlist(corr)),
         df= t.test(unlist(corr), mu = 0.5)$parameter,
         t_value = t.test(unlist(corr), mu = 0.5)$statistic,
         p = t.test(unlist(corr), mu = 0.5)$p.value,
         d = cohensD(unlist(corr), mu = 0.5))%>%
  select(age,phase,type,m,sd,t_value,df,p,d) 

iac <- data %>%
  group_by(study,phase,age, type, id) %>%
  arrange(study,phase,age, type) %>%
  summarise(corr = mean(corr)) %>%
  summarise(below=length(corr[corr<=0]),above=length(corr[corr>=1]))

# Joining tables with group level information and individual level information
ts <- t
ts$below <- iac$below
ts$above <- iac$above

# Overview
ts %>%
  knitr::kable(digits = 3)

```

# Study 1

For all models we used a maximal random effect structure (`(trial|id)`). If models did not converge, first correlation between random effects was removed and then the random slope.

```{r study 1}
### control element used for all models
contr=glmerControl(optimizer="bobyqa", optCtrl=list(maxfun=100000000))

### comparing positive and negative points in study 1
pn <- data %>%
  filter(unique_group == "pointA22neg" | unique_group =="pointA22apos") %>%
  mutate(trial=scale(trial))

### model
mpn <- glmer(corr~ unique_group+ trial+
	(trial|id),
	data=pn, family=binomial, control = contr)

### model summary and p-values
summary (mpn)
drop1(mpn, test = "Chi")

```

# Study 2

For the following models, we used all point data with negative facial expression. Models represent different assumptions about developmental trajectories.
Quadratic model did not converge with maximal random effect structure, therefore correlation between random effects is removed.
```{r study 2}
### data
np <- data %>%
  filter(phase == "point" & type =="neg") %>%
  mutate(trial=scale(trial, center = TRUE, scale=TRUE), 
         unique_group = scale(as.numeric(unique_group), center = TRUE, scale=TRUE),
         aged = scale(as.numeric(aged), center = TRUE, scale=TRUE))

### model assuming linear developmental trajectory
l <- glmer(corr~ aged+ trial+
	(trial|id),
	data=np, family=binomial, control = contr)

### model summary and p-values
summary (l)
drop1(l, test = "Chi")

### model assuming quadratic (u-shaped) developmental trajectory
q <- glmer(corr~ aged+ I(aged^2)+ trial+
	(trial|id),
	data=np, family=binomial, control = contr)

### comparing linear and quadratic model 
anova(l,q,test ="Chi")

### model summary and p-values
summary (q)
drop1(q, test = "Chi")


### model with age as categorical predictor
### data
np2 <- data %>%
  filter(phase == "point" & type =="neg") %>%
  mutate(unique_group = relevel(unique_group, ref = "pointA34neg")) %>%
  mutate(trial=scale(trial, center = TRUE, scale=TRUE))

### model
l2 <- glmer(corr~ unique_group + trial+
	(trial|id),
	data=np2, family=binomial, control = contr)

### model summary
summary (l2)

```

# Study 3

Here we analysed all data from peek conditions. Again, different models represent different assumptions about developmental trajectories.
```{r study 3}
## comparing performance across age
npe <- data %>%
  filter(phase == "peek" ) %>%
  mutate(trial=scale(trial), unique_group = scale(as.numeric(unique_group)))

### model assuming linear developmental trajectory
lp <- glmer(corr~ unique_group+ trial+
	(trial|id),
	data=npe, family=binomial, control = contr)

### model summary and p-values
summary (lp)
drop1(lp, test = "Chi")

### model assuming quadratic (u-shaped) trajectory
qp <- glmer(corr~ unique_group+ I(unique_group^2)+ trial+
	(trial|id),
	data=npe, family=binomial, control = contr)

### comparing linear and quadratic model 
anova(lp,qp,test ="Chi")


```

# Plots

## Overview of results

```{r plot results, fig.cap="Proportion of correct response per age group and study. Colored dots show means, error bars represent 95% confidence intervals around the mean based on non-parametric bootstraps. Dotted line indicated performance expected by chance. Black circles show individual data with size of the dot corresponding to the number of participants with that performance. Smoothed density distributions are shown in grey."}
ms <- data %>%
  group_by(unique_group,study,type, id) %>%
  summarise(correct = mean(corr)) %>%
  multi_boot_standard(col = "correct")

props <- data %>%
  group_by(id, type,unique_group,study, id) %>%
  summarise(corr = mean(corr))

source("geom_flat_violin.R")

ggplot()+
  geom_violin(data = props, scale = "width",adjust = 1, trim = F, aes(x = unique_group, y = corr), fill = "grey", alpha = .2)+
  geom_dotplot(data = props,aes(x = unique_group, y = corr),fill = "grey", binaxis = "y", stackdir = "centerwhole", binpositions="all", binwidth = .025, alpha = .5)+
  geom_hline(yintercept = 0.5, size = 0.5, lty=2)+
  scale_size_continuous(range = c(3,8))+
  guides(size=FALSE)+
  ylim(-0.05,1.05)+
  labs(x="Age in Month",y="Proportion Correct")+
  scale_colour_manual(name="Facial Expression",
                      labels=c("Positive", "Negative"), values=c("#859900", "#dc322f"))+
  guides(colour = guide_legend(keywidth = 2, keyheight = 2))+
  scale_x_discrete(
    breaks = c("pointA22apos", "pointA22neg","pointA28neg","pointA34neg","pointA40neg","pointA46neg","peekA22neg","peekA28neg","peekA34neg"),
    labels= c("22", "22","28","34","40","46","22","28","34"))+
  facet_grid(.~ study, scales = "free_x", space = "free_x",
          labeller = as_labeller(c(`1`="Study 1 - Pointing", `2`="Study 2 - Pointing", `3`="Study 3 - Peeking")))+
  geom_point(data=ms,aes(x = unique_group, y = mean,colour = factor(type)), pch = 5, stroke = 1, size = 2, stat="identity") + 
  geom_linerange(data = ms, aes(x = unique_group,colour = factor(type),y = mean, ymin = ci_lower, ymax = ci_upper), size = .75) +
  theme_few()
```

```{r}
#ggsave("plots/data.pdf", width = 20, height = 7, scale = 1, dpi = 2000, units = "cm")
```


## Results by trial

```{r plot trial, fig.cap="Proportion correct choice by age group and trial for conditions with negative facial expressions. Regression lines show smoothed conditional means with 95% confidence intervals. Forty- and 46-month-olds were not tested in peeking because performance was above chance at 34 month."}
ts <-  data %>%
  filter(type =="neg") %>%
  mutate(phase = ifelse(phase == "peek", "Peeking","Pointing")) %>%
  mutate(phase = as.factor(phase)) %>%
  mutate(age = recode(age, "A22"="22mo","A28"="28mo","A34"="34mo","A40"="40mo","A46"="46mo")) %>%
  mutate(phase = relevel(phase, ref = "Pointing")) %>%
  group_by(trial,age,phase,unique_group)
  
 
ts_mean <- ts %>%
  summarise(corr = mean(corr))

ggplot(ts, 
       aes(x = trial, y = corr))+
  geom_point(data = ts_mean, aes(colour=phase), size = 2, stroke = 1, pch = 21)+
  geom_smooth(method = "glm", method.args = list(family = "binomial"), se = T , level = 0.95, aes(colour=phase, fill = phase))+
  #g#ides(fill=FALSE)+
  facet_grid(.~ age, drop = T)+
  #scale_color_discrete(name="Age in month", labels=c("22","28","34","40","46"))+
  #guides(color = F)+
  ylim(0,1.05)+
  geom_hline(yintercept = 0.5, size = 1, lty=2)+
  scale_x_continuous(breaks=c(1:8), labels = c(1:8))+
  labs(x="Trial",y="Proportion Correct")+
  theme_few()+
  scale_fill_ptol(name = "Social Cue")+
  scale_color_ptol(name = "Social Cue" )


```

```{r}
ggsave("plots/trial_data.pdf", width = 20, height = 7, scale = 1, dpi = 2000, units = "cm")
```

## Different developmental trajectories

```{r}
age_continuous <-  data %>%
  filter(type =="neg", phase == "point") %>%
  group_by(aged, id) %>%
  summarise(corr = mean(corr))
  

ggplot(age_continuous, 
       aes(x = aged, y = corr))+
  geom_point( size = 1)+
  geom_smooth(method = "lm", formula = y ~ poly(x,2), se = T , level = 0.95)+
  geom_smooth(method = "lm", formula = y ~ x, se = T , level = 0.95)+
  guides(fill=FALSE)+
  #facet_grid(phase ~ age, drop = T)+
  #scale_color_discrete(name="Age in days")+
  guides(color = F)+
  ylim(0,1.05)+
  geom_hline(yintercept = 0.5, size = 1, lty=2)+
  scale_x_continuous(breaks=c(1:8), labels = c(1:8))+
  labs(x="Age in days",y="Proportion Correct")+
  theme_few()
```

